---
title: "Willamette Falls Fish Passage"
author: "Kristin Gill, Julianna Renzi, AJ Zekanoski"
date: "2/09/2021"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
```

```{r}
# reading in the data
fish_passage <- read_csv(here("data", "willamette_fish_passage.csv")) %>% 
  clean_names() %>% 
  mutate(date_new = mdy(date))
```

# Overview {.tabset}
- An engaging image (with caption, including photo credit) that is relevant to the data set

- A brief summary (3 - 4 sentences) of the dataset, and what is included in this “report”

- A map of the fish ladder location (you can make this in R on your own, or include an existing map appropriately licensed, with attribution)

- A professionally formatted data citation

- Remember: All code that you used to wrangle the data and prepare the graphs should be available to see if we click on the Code buttons. 

## Tab 1
- A finalized, static graph of adult passage for coho, jack coho, and steelhead salmon (you decide if these all occupy the same panel, or if they are in separate panels). Add a figure caption. Replace NA values with zero. 

- 2 - 3 bulletpoints summarizing major patterns / notable trends in the salmon passage time series data

```{r}
tab_1 <- fish_passage %>% 
  select(coho, jack_coho, steelhead, date_new, project, temp_c)
```

## Tab 2
## Tab 3
